---
output: github_document
---

```{r echo=FALSE, results = 'asis'}
pkg <- 'arulesNBMiner'

source("https://raw.githubusercontent.com/mhahsler/pkg_helpers/main/pkg_helpers.R")
pkg_title(pkg)
```

This R package extends [`arules`](https://michael.hahsler.net/arules) with
NBMiner, an implementation of the model-based mining algorithm for
NB-frequent itemsets described in Michael Hahsler's paper, ["A model-based
frequency constraint for mining associations from transaction
data."](https://dx.doi.org/10.1007/s10618-005-0026-2) _Data Mining and
Knowledge Discovery,_ 13(2):137-166, September 2006.

This algorithm dynamically chooses support for itemsets based on the deviation 
from a fitted independence model. NBMiner is better at suppressing spurious 
patters and finds interesting patterns with lower support compared to the traditional
minimum support-based algorithms.

```{r echo=FALSE, results = 'asis'}
pkg_install(pkg)
```

## Usage

Estimate negative binomial distribution (NBD) model parameters for the arules Groceries data set.
```{r NB_estimation}
library("arulesNBMiner")
data("Groceries")

param <- NBMinerParameters(Groceries, trim = 0.1) 
```
Note that the model shows independence but the dataset contains a significant amount
of correlated items shown by the larger cumulative frequency.


Mine NB-frequent itemsets
```{r}
itemsets_NB <- NBMiner(Groceries, parameter = param, minlen = 2L)
itemsets_NB
```

Add support and inspect some NB-frequent itemsets with the highest precision.

```{r}
quality(itemsets_NB) <- cbind(quality(itemsets_NB), 
                              support = interestMeasure(itemsets_NB, "support",
                                                        transactions = Groceries))

inspect(head(itemsets_NB, by = "precision"))
```

A precision close to 1 means that there is very little chance that 
they are spurious. We see that some very low support itemsets are under the 
top NB-frequent itemsets.


## References

* Michael Hahsler, [A
model-based frequency constraint for mining associations from
transaction data.](https://dx.doi.org/10.1007/s10618-005-0026-2)
_Data Mining and Knowledge Discovery,_ 13(2):137-166,
September 2006. 
[Preprint on ArXiv](https://doi.org/10.48550/arXiv.0803.3224)
* Michael Hahsler, Sudheer Chelluboina, Kurt Hornik, and Christian Buchta. [The arules R-package ecosystem: Analyzing interesting patterns from large transaction datasets.](https://jmlr.csail.mit.edu/papers/v12/hahsler11a.html) _Journal of Machine Learning Research,_ 12:1977-1981, 2011.
