---
output: github_document
---

```{r echo=FALSE, results = 'asis'}
pkg <- 'arulesNBMiner'

source("https://raw.githubusercontent.com/mhahsler/pkg_helpers/main/pkg_helpers.R")
pkg_title(pkg)
```

This R package 
extends package [arules](https://github.com/mhahsler/arules) with
NBMiner, an implementation of the model-based mining algorithm 
for mining NB-frequent itemsets presented in "Michael Hahsler. [A
model-based frequency constraint for mining associations from
transaction data.](https://dx.doi.org/10.1007/s10618-005-0026-2) 
_Data Mining and Knowledge Discovery,_ 13(2):137-166, September 2006." 

In addition an extension for NB-precise rules is implemented. 


```{r echo=FALSE, results = 'asis'}
pkg_install(pkg)
```

## Usage

Estimate NBD model parameters for the Agrawal data set.
```{r}
library("arulesNBMiner")
data("Agrawal")

param <- NBMinerParameters(Agrawal.db, pi = 0.99, theta = 0.5, 
                           maxlen = 5, minlen = 1, trim = 0, 
                           verbose = TRUE, plot = TRUE) 
```


Mine NB-frequent itemsets
```{r}
itemsets_NB <- NBMiner(Agrawal.db, parameter = param, 
     control = list(verb = TRUE, debug = FALSE))
```

Inspect some itemsets with the highest precision.

```{r}
inspect(head(itemsets_NB, by = "precision"))
```


## References

* Michael Hahsler, [A
model-based frequency constraint for mining associations from
transaction data.](https://dx.doi.org/10.1007/s10618-005-0026-2)
_Data Mining and Knowledge Discovery,_ 13(2):137-166,
September 2006. 
[Free preprint](https://doi.org/10.48550/arXiv.0803.3224)
* Michael Hahsler, Sudheer Chelluboina, Kurt Hornik, and Christian Buchta. [The arules R-package ecosystem: Analyzing interesting patterns from large transaction datasets.](https://jmlr.csail.mit.edu/papers/v12/hahsler11a.html) _Journal of Machine Learning Research,_ 12:1977-1981, 2011.

