---
title: "Getting started with arulesNBMiner"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with arulesNBMiner}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`arulesNBMiner` extends the `arules` package with NBMiner, which mines
negative-binomial (NB) frequent itemsets and NB-precise association rules.
This vignette demonstrates the workflow with the package's bundled Agrawal
transactions. NBMiner uses a model-based frequency constraint described by
Hahsler (2006). A single minimum-support threshold can miss associations
involving infrequent items and tends to favor short itemsets. The model-based
constraint adapts its frequency threshold to each itemset extension, using an
estimated baseline for how often independent items would co-occur.

## Load the data

Load the package and its example data. `Agrawal.db` is a `transactions`
object, the input format used by `arulesNBMiner`.

```{r data}
library(arulesNBMiner)
data("Agrawal")
summary(Agrawal.db)
```

## Fit the baseline model

The paper models each item's frequency as a Poisson count with an item
specific rate, and models variation in those rates with a Gamma distribution.
Mixing the Poisson counts over the Gamma rates gives a negative binomial
distribution. This flexible baseline captures the skewed frequency
distributions common in transaction data: a few items occur often, while many
occur rarely.

`NBMinerParameters()` estimates this global model from the item frequencies.
Items that never occur are absent from the transactions, so the estimator
uses an expectation-maximization procedure to estimate the unobserved zero
frequency class. The `trim` argument can exclude the most frequent items when
they are outliers that would distort the fit. Inspect the diagnostic plot
before choosing a nonzero trim value.

```{r parameters}
param <- NBMinerParameters(Agrawal.db, 
                           pi = 0.99, 
                           theta = 0.5, 
                           maxlen = 5, 
                           minlen = 1,
                           trim = 0, 
                           verbose = TRUE, 
                           plot = TRUE)

param
```

The figure compares the observed cumulative item frequencies with frequencies
expected under the fitted baseline. The `a` estimate is scaled per incidence
for NBMiner, as described in the paper.


## Mine NB-frequent itemsets

For an itemset already found, NBMiner considers adding each possible item. It
fits the baseline to the transactions containing that itemset and predicts
how many extensions would reach each frequency by chance. The `pi` parameter
sets the minimum predicted precision for accepting extensions. Here,
precision means the predicted proportion of accepted extensions that are
true associations under the model; it is not a guarantee of the actual
precision for every data set.

The `theta` parameter controls search pruning. For a larger itemset, it sets
the required fraction of its immediate subsets that must support the itemset
as an NB-frequent pattern. A value of 1 is most restrictive; 0 relaxes this
condition. The intermediate value 0.5 used here balances pruning with the
chance of retaining associations whose items have different frequencies.
`maxlen` limits the longest itemset considered, and `minlen` can set a
minimum length for returned patterns. The `control` list accepts `verbose`
and `debug` logical options. 

Results include model-estimated precision in
the `precision` quality column. For comparison, we add the support measure.


```{r itemsets}
itemsets <- NBMiner(
  Agrawal.db,
  parameter = param,
  control = list(verbose = FALSE, debug = FALSE)
)

quality(itemsets) <- cbind(quality(itemsets),
                           support = interestMeasure(itemsets, "support", 
                                           transactions = Agrawal.db))

inspect(head(sort(itemsets, by = "support"), n = 10))
```

## Mine NB-precise rules

NB-precise rules are rules created from NB-frequent itemsets. 
Rules instead of itemsets can be mined by changing the parameter element `rules` 
to `TRUE`.

For comparison, we add the standard measures support, confidence, and lift.

```{r rules}
param@rules <- TRUE

rules <- NBMiner(Agrawal.db, parameter = param)

quality(rules) <- cbind(quality(rules),
                           interestMeasure(rules, c("support", "confidence", "lift"), 
                                           transactions = Agrawal.db))

inspect(head(sort(rules, by = "precision"), n = 10))
```

The rules are very unlikely to random noise in the data.



## References

Hahsler, M. (2006). [A model-based frequency constraint for mining
associations from transaction data](https://doi.org/10.1007/s10618-005-0026-2).
*Data Mining and Knowledge Discovery*, 13(2), 137–166. [Full text on
arXiv](https://arxiv.org/pdf/0803.3224).
